\chapter{Requirements}
\label{chap:requirements}

In any development process, a clear and well-defined set of requirements is critical
to ensuring the success of the project. Requirements serve as a foundation for design, implementation,
verification, and validation activities throughout the software development lifecycle.
Without a structured and comprehensive understanding of what the system should do and how it 
should perform, there is risk of misalignment, scope creep, technical debt and ultimately, 
project failure. This chapter outlines the functional and non-functional requirements for the 
system. Functional requirements describe specific behaviors or functions in the system, detailing
core capabilities that directly fulfill the intended purpose of the software.
Non-functional requirements, on the other hand, define quality attributes that the system
should exhibit. These include performance, scalability and usability among others. While
non-functional requirements may not be directly observable through the system's interface, they
still have a significant impact on the user experience and overall system effectiveness.
To support effective planning and delivery, the requirements in this chapter are categorized 
using the MoSCoW method:
\begin{itemize}
    \item Must Have - Essential features that without the system cannot function.
    \item Should Have - Important features that add significant value but are not critical 
    for a minimum viable product (MVP).
    \item Could Have - Desirable features that may be included if time and resources allow for it.
    \item Won't Have - Requirements that explicitly are deferred for future development or out of scope.
\end{itemize}

\input{chapters/requirements/sec-functional.tex}
\input{chapters/requirements/sec-non-functional.tex}
