\section{Functional Requirements}
\label{sec:functional-requirements}

\newcounter{requirement}
\newcounter{subrequirement}
\newcommand\rownumber{\arabic{requirement}}
\newcommand\subrownumber[1]{%
  \stepcounter{subrequirement}%
  \arabic{requirement}.\arabic{subrequirement}%
}
\newcommand\resetrownumbers{\setcounter{requirement}{0}\setcounter{subrequirement}{0}}

\begin{xltabular}{\textwidth}{
  >{\raggedright\arraybackslash}m{1cm} |
  >{\raggedright\arraybackslash}X |
  >{\raggedright\arraybackslash}X |
  >{\centering\arraybackslash}m{2cm}
  }
  \textbf{ID}                                              & \textbf{Title}          & \textbf{Description}                                                                             & \textbf{MoSCoW} \\\hline
  \stepcounter{requirement}F\rownumber                     & Battle Environment      & Turn-based battle simulator that can take inputs from an agent.                                  & M               \\\hline
  F\subrownumber{\rownumber}                               & Gen 9 Mechanics         & The environment will be based on the combat mechanics currently available in Generation 9 games. & M               \\\hline
  F\subrownumber{\rownumber}                               & Game Specific Mechanics & Game specific mechanics such as Mega Evolution, Z-moves, Terastalization and Dynamaxing.         & W               \\\hline
  F\subrownumber{\rownumber}                               & Uncommon status conditions & Certain status conditions are very rare and don't regularly affect the outcome of a battle.         & C               \\\hline
  F\subrownumber{\rownumber}                               & Uncommon unique moves & Certain unique moves that require custom implementation and don't regularly appear in the games, such as moves exclusive to a single Pokemon species.         & C               \\\hline
  F\subrownumber{\rownumber}                               & Held Items              & The environment supports a subset of held items.                                                 & M               \\\hline
  F\subrownumber{\rownumber}                               & Single Battles          & The environment must support single battles with 1v1 combat.                                     & M               \\\hline
  F\subrownumber{\rownumber}\setcounter{subrequirement}{0} & Double Battles          & The environment must support double battles with 2v2 combat.                                     & S               \\\hline
  \stepcounter{requirement}F\rownumber                     & Trainable Agent         & An agent that can be trained in the battle environment.                                          & M               \\\hline
  \stepcounter{requirement}F\rownumber                     & Configurable Teams      & The sets of Pokemon used in the environment can be changed.                                      & S               \\\hline
  \stepcounter{requirement}F\rownumber                     & Team Builder            & Build a team using a GUI for the agent(s) to use.                                                & C               \\\hline
  \stepcounter{requirement}F\rownumber                     & GUI                     & A GUI for viewing the agents actions and watching combat logs in real time.                      & C               \\\hline
  \caption{Functional Requirements}
  \label{tab:functional-requirements}
\end{xltabular}
